---
title: Empirical Exercises 4.2
author: Zheng Tian
output: 
  html_document: default
---

This file include answers and R codes for completing Empirical Exercise 4.2 in Introduction to Econometrics (3rd edition) by Stock and Watson.

Reading the Data
================

The first step is to read the data file into R. The data files for this problem are `TeachingRatings.dta` and `TeachingRatings.xls`, accompanied by a descriptive file `TeachingRatings_Description.pdf`.

-   Read the STATA file

```{r}
library(foreign)
teachingdata <- read.dta("TeachingRatings.dta")
```

- Upon reading the data, we can take a glimpse on the data. Use `head` or `tail` to look at the first or last few observations

```{r}
head(teachingdata)
```


Summary Statistics
==================

We get the summary statistics of the variables used in the analysis, which is `course_eval` and `beauty`

```{r}
df <- teachingdata[c("course_eval", "beauty")]
sumdf <- summary(df); sumdf
```

We can create a table that looks professional using `stargazer()`.
```{r, results='asis'}
library(stargazer)
stargazer(df, type = "html", header = FALSE,
          title = "Summary Statistics", label = "tab:sum-stats")
```


Scatterplot
===========

We can make scatterplot using the `plot` function.

```{r}
  teaching.formula <- course_eval ~ beauty
  plot(teaching.formula, data = teachingdata,
     main = "The Scatterplot of Course Evaluation on Professor's Beauty",
     xlab="Beauty", ylab = "Course evaluation", col = "blue")
```


Regression
==========

Now let's estimate the regression model. The results is reported in Table

```{r, tidy=TRUE, results='asis'}
# run a regression of course evaluation on professor's beauty
teaching.ols <- lm(teaching.formula, data = teachingdata)

# create the table
stargazer(teaching.ols, type = "html",
  covariate.labels = c("Beauty"),
  dep.var.labels = c("Course Evaluations"),
  title = "The OLS Estimation of the Regression of Course Evaluation on Beauty",
  label = "tab:ols-1", single.row = TRUE, omit.stat = c("adj.rsq", "f")
)
```



Answers to the questions
========================

a.  
The scatterplot is the figure shown above. There appears to be a weak positive relationship between course evaluation and the beauty index.

b.  
The estimation results are reported in the table above .

We use the following code the get the answers to (b) and (c).

```{r}
beauty.watson <- mean(teachingdata$beauty)
beauty.stock <- mean(teachingdata$beauty) + sd(teachingdata$beauty)
ave.courseval <- mean(teachingdata$course_eval)

# do prediction step by step
b0 <- teaching.ols$coef[1]
b1 <- teaching.ols$coef[2]
courseval.predict <- b0 + b1 * c(beauty.watson, beauty.stock)
names(courseval.predict) <- c("waston", "stock")
```

The slope is `r round(b1, 4)` and the intercept is `r round(b0, 3)`. The sample mean of course evaluation is `r round(ave.courseval, 3)`, which coincides with the slope because the sample mean of *Beauty* is `r round(beauty.watson, 4)`.

c.  
The beauty indices for Professors Stock and Watson are `r round(beauty.stock, 4)` (one standard deviation) and `r round(beauty.watson, 4)` (sample average). Thus, the predicted course evaluations for Professors Stock and Watson are `r round(courseval.predict[2], 4)` and `r round(courseval.predict[1], 4)`, respectively.

d.

The following code is to answer (d)

```{r}
beauty.sd <- sd(teachingdata$beauty)
courseval.sd <- sd(teachingdata$course_eval)
delta.courseval <- b1 * beauty.sd
```

The standard deviation of course evaluation is `r round(courseval.sd, 4)`, and the standard deviation of beauty is `r round(beauty.sd, 4)`. A one-standard-deviation increase in beauty is expected to increase course evaluation by `r round(delta.courseval, 4)`, or `r round(delta.courseval/courseval.sd, 2)` of standard deviation of course evaluations. The effect is small.

e. 

```{r}
rsq <- summary(teaching.ols)$r.squared
```


The regression R<sup>2</sup> is `r round(rsq, 4)`, so that *Beauty* explains only `r round(rsq, 3) * 100` percent of the variance in course evaluations.


