---
title: "Answers to Empirical Exercise 3.1 with R"
output:
  html_document: default
  pdf_document: default
---

Now we can show how to use R to do some statistical analysis. This
demonstration answers the questions of Empirical Exercise 3.1 at the end
of Chapter 3. Furthermore, we carry out this exercise in the format of
**reproducible research**. That means, we should accomplish they
following tasks in answering the problem:

1.  using R to compute the statistics asked in the questions
2.  including R code and the results of running the code in the answer,
    and
3.  describing our work and answers in plain language along with code
    and numerical answers.

# A description of the problem

Empirical exercise 3.1 concerns the relationship between average
earnings and education levels, using the data set from the 1992 and 2008
Current Population Survey (CPS). Specifically, we want to see whether
the average hourly earnings (`ahe`) are different between workers with a
bachelor degree and those with only high school diploma (`bachelor`).

# Answers to the questions

## Question (a)

Compute the sample mean for average hourly earnings (`ahe`) in 1992 and\
in 2008. Construct a 95% confidence interval for the population means\
for `ahe` in 1992 and 2008 and the change between 1992 and 2008

### Read the data

The first thing first is of course read the data correctly from the
Stata file [file:../data/cps92\_08.dta](data/cps92_08.dta), which can be
read by the function `read.dta()` in the package of `foreign`.

``` {r, echo=TRUE, results='asis', include=TRUE}
  library(foreign)
  cpsdat <- read.dta("cps92_08.dta")
  head(cpsdat)
```


### Calculate the sample means of average hourly earnings in 1992 and 2008

There are many ways to compute the sample means in 1992 and 2008,
respectively. First, to make you more familiar with the R language, we
compute them in a very basic way. Then, we show how to get the same
results with some powerful functions.

```{r echo=TRUE, include=TRUE, results='hide'}
  # extract the data for average hourly earnings in 1992 and 2008
  ahe.92 <- cpsdat$ahe[cpsdat$year == 1992]
  ahe.08 <- cpsdat$ahe[cpsdat$year == 2008]
  mean.ahe.92 <- mean(ahe.92); mean.ahe.92
  mean.ahe.08 <- mean(ahe.08); mean.ahe.08
```

The average hourly earnings are `r round(mean.ahe.92, 2)` dollars in 1992 and
`r round(mean.ahe.08, 2)` dollars in 2008.

### Construct the confidence intervals

Recall that a 95% confidence interval for the population mean can be
constructed as $\overline{Y} \pm 1.96 SE(\overline{Y})$ and
$SE(\overline{Y})$ is computed as $s_Y / \sqrt{n}$.

```{r echo=TRUE}
# the sample variance
sd.ahe.92 <- sd(ahe.92)
sd.ahe.08 <- sd(ahe.08)

n.92 <- length(ahe.92)
n.08 <- length(ahe.08)

# the standard error
se.ahe.92 <- sd.ahe.92 / sqrt(n.92)
se.ahe.08 <- sd.ahe.08 / sqrt(n.08)

# 95% confidence interval
# the 95% critical value from a normal distribution
cv.95 <- qnorm(0.975)

lower.lim.92 <- mean.ahe.92 - cv.95 * se.ahe.92
lower.lim.08 <- mean.ahe.08 - cv.95 * se.ahe.08

upper.lim.92 <- mean.ahe.92 + cv.95 * se.ahe.92
upper.lim.08 <- mean.ahe.08 + cv.95 * se.ahe.08
```

The 95% confidence interval for `ahe` in 1992 is
(`r round(lower.lim.92, 2)`, `r round(upper.lim.92, 2)`),
and that in 2008 is (`r round(lower.lim.08, 2)`, `r round(upper.lim.08, 2)`).

### Alternative methods to calculate the sample means and confidence intervals

In the above example, to compute the sample averages in 1992 and 2008,
we write code separately for each year, which can be done more easily in
R.

We can compute the averages for each year using the function
`aggregate()`, which splits the whole data base into two parts by the
values of `year`. Then, for each part we compute the average by
specifying the argument `FUN` to be `mean`, i.e., specifying the
function to be used for each part as the `mean()` function. Also, in
this case, we use `~` to specify a **formula** that means that we split
`ahe` by `year`.

```{r echo=TRUE, results="asis", include=TRUE}
  # Use aggregate() to compute the means in both years
  ahe.means <- aggregate(ahe ~ year, FUN = mean, data = cpsdat)
  ahe.means
```

The confidence interval can be extracted from the results of the
`t.test()` function, which is a list.

```{r echo=TRUE, results='asis', include=TRUE}
  # t test for ahe in 1992
  t.ahe.92 <- t.test(ahe.92); t.ahe.92$conf.int
  # t test for ahe in 2008
  t.ahe.08 <- t.test(ahe.08); t.ahe.08$conf.int
  # test for the change between 1992 and 2008
  t.ahe.diff <- t.test(ahe.08, ahe.92); t.ahe.diff
```

The confidence interval of the change in average hourly earnings between
1992 and 2008 is (`r round(t.ahe.diff$conf.int[1], 2)`,
`r round(t.ahe.diff$conf.int[2], 2)`).

Question (b)
------------

Now we need to adjust the average hourly earnings in the 1992 dollars to
the 2008 dollars with the inflation rate, computed as `CPI2008/CPI1992`.

```{r echo=TRUE, results='hide', include=TRUE}
  # CPI in 1992 and 2008
  cpi.92 <- 140.3
  cpi.08 <- 215.2
  # Inflation adjustment
  inflator <- cpi.08 / cpi.92
  cpsdat$ahe.adj <- with(cpsdat, ifelse(year == 1992, ahe * inflator, ahe))
```

In the code block above, we first use the function `with()` to attach
the data frame `cpsdat` within its own environment so that when we refer
to variables in `cpsdat`, such as `ahe` and `year`, we do not need to
write `cpsdat$` and every time we use its variables.

The function `ifesle()` set the values of `ahe` based on the condition
`year == 1992`. If the condition is true, we do `ahe *
inflator`; if not, leave `ahe` as it is.

Then we repeat what we've done in Question (a) with the
inflation-adjusted earnings in 1992.

```{r echo=TRUE}
  ahe.92.adj <- with(cpsdat, ahe.adj[year == 1992])
  mean.ahe.92.adj <- mean(ahe.92.adj)
  t.ahe.92.adj <- t.test(ahe.92.adj)
  t.ahe.diff.adj <- t.test(ahe.08, ahe.92.adj)
```

-   The sample average of the inflation-adjusted earnings in 1992 is
    `r round(mean.ahe.92.adj, 2)` in
    the 2008 dollars.

-   The confidence interval for the inflation-adjusted average hourly
    earnings in 1992 is
    (`r paste0(round(t.ahe.92.adj$conf.int, 2), collapse = ", ")`).

-   The confidence interval for the change between 1992 and 2008 is
    (`r paste0(round(t.ahe.diff.adj$conf.int, 2), collapse = ", ")`).

Question (c)
------------

If we are interested in the change in workers' purchasing power, the
results with the inflation-adjusted earnings should be used in
comparison.

Question (d)
------------

Now let's compute the average earnings for high school graduates and
college graduates with the 2008 data. First thing to do is to select the
2008 data from `cpsdat` using the function `subset()`

```{r echo=TRUE}
  # select data in 2008
  cps08 <- subset(cpsdat, year == 2008, select = c(year, ahe, bachelor))

  # calculate means
  ahe.educ.08 <- aggregate(ahe ~ bachelor, FUN = mean, data = cps08)

  # select ahe and filter by bachelor
  ahe.high.08 <- with(cps08, ahe[bachelor == 0])
  ahe.bach.08 <- with(cps08, ahe[bachelor == 1])

  # construct confindence interval
  t.ahe.high.08 <- t.test(ahe.high.08)
  t.ahe.bach.08 <- t.test(ahe.bach.08)
  t.ahe.gap.08  <- t.test(ahe.bach.08, ahe.high.08)
```

-   The mean of the average hourly earnings of high school graduates in
    2008 is `r round(ahe.educ.08$ahe[1], 2)` dollars with the 95% confidence interval
    (`r paste0(round(t.ahe.high.08$conf.int, 2), collapse = ", ")`).

-   The mean of the average hourly earnings of college graduates is
    `r round(ahe.educ.08$ahe[2], 2)` dollars with the 95% confidence interval
    (`r paste0(round(t.ahe.bach.08$conf.int, 2), collapse=", ")`).

-   The 95% confidence interval of the gap in earnings between the two
    groups is
    (`r paste0(round(t.ahe.gap.08$conf.int, 2), collapse=", ")`).


We can create a boxplot to compare the means and confidence intervals of
average hourly earnings between high school graduates and college
graduates.

```{r echo=TRUE, fig.width=6, fig.height=6, fig.align='center'}
boxplot(ahe ~ bachelor, data = cps08,
        main = "Average Hourly Earnings by Education",
        col = c("red", "orange"),
        xlab = "Bachelor degres = 1, high school = 0",
        ylab = "US$ 2008")
```

We leave Question (e)-(g) to students as exercises.
