---
title: "Homework_2"
author: "Wang Tianqi & Hao Jinsong"
date: "2017Äê4ÔÂ3ÈÕ"
output:
  html_document: default
  pdf_document: default
---

#Homework_2 Answer
##Empirical Exercise
###4.2a
```{r}
library(foreign)
professordata <- read.dta("TeachingRatings.dta")
head(professordata[c("beauty", "course_eval")])

df <- professordata[c("beauty","course_eval")]
plot(df$beauty, df$course_eval,
     main = "Course evaluations on the Professor's beauty", 
     xlab = "Beauty", 
     ylab = "Course_eval",
     col = "blue", pch =16, cex = 0.7, bty = "l")
cor(df$beauty,df$course_eval)
```
###4.2b
Course_eval = 3.99827 + 0.133Beauty

The sample mean of Beauty is 0. So, the estimated intercept is equal to the sample mean of Course_Eval
```{r}
mod1 <- lm(course_eval ~ beauty, data = df)
summary(mod1)
plot(df$beauty, df$course_eval,
     main = "Course evaluations on the Professor's beauty", 
     xlab = "Beauty", 
     ylab = "Course_eval",
     col = "blue", pch =16, cex = 0.7, bty = "l")
intercept <- coef(mod1)[1]
slope <- coef(mod1)[2]
abline(intercept, slope, col="red")
texteq <- paste("Course_eval = ", round(intercept, 1), " + ",
                round(slope, 2), "Beauty", sep = "")
text(1, 3.9, texteq, cex.lab = 0.9, font.lab = 3)
```

###4.2c
```{r}
std <- sapply(df,sd)
std
coeffs <- coefficients(mod1)
#Prof.Watson 
Beauty <- mean(df$beauty)
course_eval <- coeffs[1] + coeffs[2]*Beauty 
course_eval

#Prof. Stock
Beauty <- mean(df$beauty)+sd(df$beauty)
course_eval <- coeffs[1] + coeffs[2]*Beauty 
course_eval
```

###4.2d
The estimated effect of Beauty on Course_Eval is small.
The standard deviation of beauty is 0.7886477 and the stand deviation of course evaluation is 0.5548656.
0.133 * 0.789 = 0.104937, far less than the std of Course Evaluation.

###4.2e
Multiple R-squared:  0.03574. Beauty explains a small fraction.