---
title: "Econometric Empirical Exercise"
author: "group2"
date: "2017.4.4"
output:
  html_document: default
  pdf_document: default
  word_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Probleam a.
```{r pressurea, echo=FALSE}
library(foreign)
setwd("C:/STUDY/econometrics/homework/TeachingRatings")
classdata <- read.dta("TeachingRatings.dta")
df <- classdata[c("beauty", "course_eval")]

     
plot(df$beauty, df$course_eval, col = "blue", pch =16, cex = 0.7, bty = "l",
main = "scatterplot of average course evaluations on the professor£ªs beauty",
xlab = "Beauty", ylab = "Course_eval")



```


Answer for a:
It seems that there is a weak positive relation between course evaluation and the beauty index.

#Problem b.
```{r pressureb, echo=FALSE}
setwd("C:/STUDY/econometrics/homework/TeachingRatings")
classdata <- read.dta("TeachingRatings.dta")
mod1 <- lm(course_eval ~ beauty, data = df)
summary(mod1)
summary(df)




```
Anwser for b:
Course_eval=3.998+0.133Beauty
The estimated intercept is 3.998.
The estimated slope is 0.133.
Because the mean of beauty is near to zero.So due to the euqation,the estimated intercept is equal to the mean of Course_Eval.
```{r pressurec, echo=FALSE}
plot(df$beauty, df$course_eval, col = "blue", pch =16, cex = 0.7, bty = "l",
xlab = "beauty", ylab = "course_eval")
intercept <- coef(mod1)[1]
slope <- coef(mod1)[2]
abline(intercept, slope, col="red")
texteq <- paste("Course_eval = ", round(intercept, 1), " + ",
round(slope, 2), "beauty", sep = "")
text(1.5, 4.4, texteq, cex.lab = 0.9, font.lab = 3)

```
#Problem c.
```{r pressured, echo=FALSE}
std <- sapply(df, sd)
print(std)

```
Answer for c:
The standard deviation of Beauty is 0.789. 
Thus,
Professor Watson£ªs predicted course evaluations = 4.00;
Professor Stock£ªs predicted course evaluations =4.00+0.133¨G1¨G0.789=4.105.

#Problem d:
Answer for d:
The standard deviation of course evaluations is 0.55 and the standard deviation of beauty is
0.789. 
A one standard deviation increase in beauty is expected to increase course evaluation by 0.105.
It's much smaller than a standard deviation of course evaluations. Thus,the effect is small.

#Problem e:
Answer for e:
No,  R-squared is 0.03574,much less than 1, so that Beauty cannot explain the course evaluation efficiently.



