---
title: "A Brief R Markdown Tutorial"
author: "Zheng Tian"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown

R Markdown provides an authoring framework for data science. You can use a single R Markdown file to both

* save and execute code
* generate high quality reports that can be shared with an audience

An R Markdown file is a plain text file that has the extension `.Rmd`. In RStudio, you can easily create a new R Markdown file through the following steps:

1. Click `File -> New File -> R Markdown`. 
2. At the left panel in the window jumping up, choose `Document`, and at the    right panel, enter the title and author name of the document. Then, choose    `HTML` as the default output format and click `OK`. 

A short R Markdown document is generated with some elements in it as a template. 

## How it works

To create an `HTML` document from a R Markdown file, click `Knit` button in RStudio. A built-in HTML browser will be invoked to have a preview of the document. You can also create a PDF file by click the little arrow beside `Knit` button, and choose `Knit to PDF`. 

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
